Verification of a TCP/IP stack

### 2021-05-21 SecDev21/paper.tex
Change title and remove reference to code rep for anonymity
fixing C to SPARK section

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}. SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification. For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack. By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library. 
\todo[inline]{pending contributions list and fixing the next paragraph.}
We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

@@

To reduce the size of the work, only few functions of interest have been
translated in SPARK. Most of the C code has been kept.
Hopefully, SPARK integrates a mechanism to interface C code. No verification are
performed on the C code, but pre- and postconditions can be attached to
subprograms and they can be called in SPARK code.
The functional specification are not proved, thus it is the role of the
programmer to ensure the correctness of the contracts.

>>>>

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}. SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification. For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack. By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library. 
\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

@@

The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work. Instead, the aim is to demonstrate how the use of SPARK can provide significant security hardening to widely used C-written libraries. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API). This mainly falls under two categories: a)  hardening the user's API to enforce its correct usage by the users, discussed in \Cref{sec:API}, and b) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}.

\todo[inline]{KG: fixed up to here}

Hopefully, SPARK integrates a mechanism to interface C code. No verification are
performed on the C code, but pre- and postconditions can be attached to
subprograms and they can be called in SPARK code.
The functional specification are not proved, thus it is the role of the
programmer to ensure the correctness of the contracts.
